---
title: "DATA-205-Capstone-Project"
author: "Fiona Caretto"
date: "4/28/2021"
output: html_document
---

# Loading in datasets

First, I read in two dataMontgomery datasets: "Moderately Priced Dwelling Units" (named as "MPDU") and "Housing Licensing and Registration" (named as "HLR").

```{r}
MPDU <- read.csv("https://data.montgomerycountymd.gov/api/views/pbh5-g4e5/rows.csv?accessType=DOWNLOAD")
HLR <- read.csv("https://data.montgomerycountymd.gov/api/views/et5s-xste/rows.csv?accessType=DOWNLOAD")
```

Next, I will read in the "Students Receiving Special Services" dataset from the Maryland State Department of Education, one for each year 2015-2019. Each of these will be called "SS+last two digits of year."

```{r}
url1 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/378"
download.file(url1, "Special_Services_2019.zip")
unzip("Special_Services_2019.zip")
SS19 <- read.csv("Special_Services_2019.csv")
```

```{r}
url2 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/341"
download.file(url2, "Special_Services_2018.zip")
unzip("Special_Services_2018.zip")
SS18 <- read.csv("Special_Services_2018.csv")
```

```{r}
url3 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/322"
download.file(url3, "Special_Services_2017.zip")
unzip("Special_Services_2017.zip")
SS17 <- read.csv("Special_Services_2017.csv")
```

```{r}
url4 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/298"
download.file(url4, "Special_Services_2016.zip")
unzip("Special_Services_2016.zip")
SS16 <- read.csv("Special_Services_2016.csv")
```

```{r}
url5 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/258"
download.file(url5, "Special_Services_2015.zip")
unzip("Special_Services_2015.zip")
SS15 <- read.csv("Special_Services_2015.csv")
```

Now, I will read in the "Adjusted Cohort Graduation Rate" datasets also from the Maryland State Department of Education, also for 2015-2019. Each of these will be named "Grad+last two digits of year."

```{r}
url6 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/390"
download.file(url6, "Cohort_Graduation_and_Dropout_Rates_2019.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2019.zip")
Grad19 <- read.csv("Cohort_Grad_Rate_2019.csv")
```

```{r}
url7 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/354"
download.file(url7, "Cohort_Graduation_and_Dropout_Rates_2018.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2018.zip")
Grad18 <- read.csv("Cohort_Grad_Rate_2018.csv")
```

```{r}
url8 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/326"
download.file(url8, "Cohort_Grad_Rate_2017.zip")
unzip("Cohort_Grad_Rate_2017.zip")
Grad17 <- read.csv("Cohort_Grad_Rate_2017.csv")
```

```{r}
url9 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/302"
download.file(url9, "Cohort_Grad_Rate_2016.zip")
unzip("Cohort_Grad_Rate_2016.zip")
Grad16 <- read.csv("Cohort_Grad_Rate_2016.csv")
```

```{r}
url10 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/262"
download.file(url10, "Cohort_Grad_Rate_2015.zip")
unzip("Cohort_Grad_Rate_2015.zip")
Grad15 <- read.csv("Cohort_Grad_Rate_2015.csv")
```

Lastly, I will read in the Low-Income Housing Tax Credit dataset from the Department of Housing and Urban Development.

```{r}
url11 <- "https://lihtc.huduser.gov/lihtcpub.zip"
download.file(url11, "lihtcpub.zip")
unzip("lihtcpub.zip")
LIHTC <- read.csv("LIHTCPUB.CSV")
```

To start preparing to merge my datasets along school names and zip codes, I am going to make two new data frames: one with all MCPS high schools and their zip codes and one with all PGCPS high schools and their zip codes.

I will install the tidyverse package in order to do this.

```{r}
install.packages("tidyverse")
library(tidyverse)
MCPS <- data.frame(
  School.Name = c("Stephen Knolls School", "Longview School", "Montgomery Blair High", "Northwest High", "Walter Johnson High", "Richard Montgomery High", "Gaithersburg High", "Clarksburg High", "Winston Churchill High", "Bethesda-Chevy Chase High", "Quince Orchard High", "Thomas S. Wootton High", "Walt Whitman High", "Wheaton High", "Sherwood High", "Paint Branch High", "John F. Kennedy High", "Albert Einstein High", "Northwood High", "Springbrook High", "James Hubert Blake High", "Watkins Mill High", "Col. Zadok Magruder High", "Rockville High", "Damascus High", "Poolesville High", "Seneca Valley High", "Alternative Programs", "John L Gildner Regional Inst for Children & Adol", "Rock Terrace School"),
  School.Zip = c(20895, 20874, 20901, 20874, 20814, 20852, 20877, 20871, 20854, 20814, 20878, 20850, 20817, 20906, 20860, 20866, 20902, 20895, 20901, 20904, 20905, 20879, 20855, 20851, 20872, 20837, 20874, 20853, 20850, 20850))
PGCPS = data.frame(
  School.Name = c("Eleanor Roosevelt High","High Point High","Bowie High","Parkdale High","Northwestern High","Dr. Henry A. Wise, Jr. High","Charles Herbert Flowers High","Duval High","Suitland High","Bladensburg High","Laurel High","Oxon Hill High","Potomac High","Frederick Douglass High","Crossland High","Gwynn Park High","Largo High","Central High","Friendly High","Fairmont Heights High","Surrattsville High","Academy of Health Sciences at PGCC","Chesapeake Math and IT Public Charter","International High school @ Largo","Northwestern Evening/Saturday High","International High School @ Langley Park","Chesapeake Math and IT South Public Charter","College Park Academy","Crossland Evening/Saturday High","Tall Oaks High","Croom High","Community Based Classrooms","C. Elizabeth Rieg Regional Center","James E. Duckworth Regional Center"),
  School.Zip = c(20770, 20705, 20715, 20737, 20782, 20772, 20774, 20706, 20746, 20710, 20707, 20745, 20745, 20772, 20748, 20613, 20774, 20747, 20744, 20785, 20735, 20774, 20707, 20772, 20782, 20710, 20772, 20737, 20748, 20721, 20623, 20710, 20716, 20705))

MCPS <- MCPS %>% add_row(School.Name = "Regional Institute for Children & Adolescents", School.Zip = 20850)
PGCPS <- PGCPS %>% add_row(School.Name = "Annapolis Road Academy", School.Zip = 20716) %>% add_row(School.Name = "Green Valley Academy at Edgar Allan Poe", School.Zip = 20746)
School_Zips <- rbind(MCPS, PGCPS)
```

I had to add some rows afterwards that I had initially, unintentionally left out of these data frames.

# Looking at the structure of the datasets

```{r}
head(HLR)
tail(HLR)
summary(HLR)
```
```{r}
head(LIHTC)
tail(LIHTC)
summary(LIHTC)
```
```{r}
head(MCPS)
tail(MCPS)
summary(MCPS)
```
```{r}
head(MPDU)
tail(MPDU)
summary(MPDU)
```
```{r}
head(PGCPS)
tail(PGCPS)
summary(PGCPS)
```
```{r}
head(School_Zips)
tail(School_Zips)
summary(School_Zips)
```
```{r}
head(Grad15)
tail(Grad15)
summary(Grad15)
```
```{r}
head(SS15)
tail(SS15)
summary(SS15)
```
#Wrangle, clean, mutate the data.

I'm going to create a new data frame out of the MPDU dataset showing just Total Units and Total MPDU, grouped by Planning Area, naming the new dataset "MPDU2."

```{r}
MPDU2 <- MPDU %>% 
select(Planning.Area, Total.Units, Total.MPDU) %>%
group_by(Planning.Area)
MPDU2
```
From this new data frame, I'm going to create a new one showing the sum of Total Units and Total MPDU, by Planning Area. First, I need to drop the NAs from the MPDU2 data frame.

```{r}
MPDU2 <- drop_na(MPDU2)
MPDU_Sums <- MPDU2 %>% 
    summarize_if(is.integer, sum)
MPDU_Sums
```

Now I will create a new data frame with variables of interest from the HLR dataset: Ownership Type, Property Zip Code, and Unit Count.

```{r}
HLR2 <- HLR %>% 
select(Ownership.Type, Property.Zip.Code, Unit.Count) %>%
group_by(Ownership.Type)
HLR2
```

Turning to my Adjusted Cohort Graduation Rates datasets, I need to do a bit of cleaning so that the columns of all these datasets match, making it easier to combine them later on.

```{r}
Grad15 <- Grad15 %>% 
  rename(Academic.Year = Class.of.Year)
Grad16 <- Grad16 %>% 
  rename(Academic.Year = Class.of.Year)
```

```{r}
Grad17 <- Grad17 %>%
  rename(LEA = LEA.Number)
Grad18 <- Grad18 %>%
  rename(LEA = LSS.Number, LEA.Name = LSS.Name)
Grad19 <- Grad19 %>%
  rename(LEA = LSS.Number, LEA.Name = LSS.Name)
```

Now, I will combine all five Adjusted Cohort Graduation Rates into one data frame.

```{r}
Total_Grad <- rbind(Grad15, Grad16, Grad17, Grad18, Grad19)
```

For my project, I am only interested in looking at the data for Montgomery County and Prince George's County, so I'm going to pull out that data.

```{r}
Total_Grad2 <- subset(Total_Grad, LEA.Name == "Montgomery" & Cohort == "4-year adjusted cohort" | LEA.Name == "Prince George's" & Cohort == "4-year adjusted cohort")
```

In the Graduation Rates column, some of the values are represented by inequalities: <= 5.00 and >= 95.00. Since I only want numbers in this column, I'm going to replace these values with 5 and 95 respectively.

```{r}
Total_Grad2[Total_Grad2== ">= 95.00"] <-  "95.00"
```

```{r}
Total_Grad2[Total_Grad2== "<= 5.00"] <-  "5.00"
```

Now, I will convert the Graduation Rate variable from characters to numerics.

```{r}
Total_Grad2$Grad.Rate <- as.numeric(Total_Grad2$Grad.Rate)
```

Taking a look at the NAs introduce by coercion.

```{r}
summary(Total_Grad2)
```

Turning to the Students Receiving Special Services datasets, I will first rename some columns so that all of the datasets match up, and then I'll create new data frames pulling in only the variables I am interested in.

```{r}
SS15_2 <- SS15 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct)
SS15_2

SS16_2 <- SS16 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct)
SS16_2

SS17_2 <- SS17 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct)
SS17_2

SS18_2 <- SS18 %>% 
  rename(LEA.Name = LSS.Name) %>%
  select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct)
SS18_2

SS19_2 <- SS19 %>% 
  rename(LEA.Name = LSS.Name) %>%
  select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct)
SS19_2
```

```{r}
Total_SS <- rbind(SS15_2, SS16_2, SS17_2, SS18_2, SS19_2)
```

Now, I will combine my Graduation Rates and Special Services data frames.

```{r}
Total_Grad_SS <- merge(Total_Grad2, Total_SS)
```

Remove Elementary and Middle Schools from this data frame.

```{r}
Total_Grad_SS <- Total_Grad_SS %>%
  filter(substr(School.Type, 1, 4) == "High")
```

There were multiple instances where the school name was slightly off ("Northwood High School" instead of "Northwood High"), so I renamed them.

```{r}
Total_Grad_SS[Total_Grad_SS == "Northwood High School"] <- "Northwood High"
```

```{r}
Total_Grad_SS[Total_Grad_SS == "Tall Oaks Vocational"] <- "Tall Oaks High"
```

```{r}
Total_Grad_SS[Total_Grad_SS == "Croom Vocational"] <- "Croom High"
```

I am now removing the rows for "All Montgomery Schools" and "All Prince George's Schools," and all the rows of schools that either don't exist anymore or ones that I couldn't find sufficient or consistent information on online.

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="All Montgomery Schools")
```

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="All Prince George's Schools")
```

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="Forestville High")
```

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="Incarcerated Youth Center (JACS)")
```

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="Margaret Brent Regional Center")
```

```{r}
Total_Grad_SS <- subset(Total_Grad_SS, School.Name!="MCPS Transitions School")
```

Now, I am creating a new column combining School Names and Academic Year so that there's a unique identifier for each row.

```{r}
Total_Grad_SS$School.By.Year <- paste(Total_Grad_SS$School.Name, Total_Grad_SS$Academic.Year)
```

Now, I will merge the Total_Grad_SS dataset with the School_Zips data frame I made of all school names and their zip codes. Now I have one full data frame with all the school data and their zip codes.

```{r}
School <- merge(Total_Grad_SS, School_Zips, by.x = "School.Name", by.y = "School.Name")
```

Turning back to my Housing Licensing and Registration data, I will now create a new data frame from the HLR2 data frame showing the total amount of units under each Ownership Type by Property Zip Code.

```{r}
HLR_zip <- aggregate(HLR2$Unit.Count, by=list(Category=HLR2$Property.Zip.Code, HLR2$Ownership.Type), FUN=sum)
HLR_zip
```
Now, for my LIHTC dataset, I am going to pull out the data only for Montgomery and Prince George's county, by extracting this data from the "fips2010" variable. This variable is the FIPS code of the public housing project, which contains the State FIPS code, County FIPS code, and Census Tract number.

```{r}
LIHTC2 <- LIHTC %>%
  filter((substr(fips2010, 1, 5) == "24031") | (substr(fips2010, 1, 5) == "24033"))
```

Now, I am preparing to merge the LIHTC data with my School data frame. First, I will rename the zip code variable of the LIHTC data frame, and then change the data type of the zip code in the School data frame to "character," to match the zip code variable type in the LIHTC data frame.

```{r}
LIHTC2 <- LIHTC2 %>% 
  rename(
   School.Name = proj_zip)
```

Oops, changed it to the wrong name.

```{r}
LIHTC2 <- LIHTC2 %>% 
  rename(
   School.Zip = School.Name
    )
```

```{r}
LIHTC3 <- aggregate(LIHTC2$li_units, by=list(Category=LIHTC2$School.Zip), FUN=sum)
LIHTC3
```
```{r}
LIHTC3 <- LIHTC3 %>% 
  rename(Total_LIunits = x)
```

```{r}
LIHTC3 <- LIHTC3 %>% 
  rename(School.Zip = Category)
```

```{r}
School$School.Zip <- as.character(School$School.Zip)
```

```{r}
School2 <- merge(School, LIHTC3, by.x = "School.Zip", by.y = "School.Zip", all = TRUE)
```

My data frames were unequal in size, so when I merged them, 15 extra rows were created so I removed them here.

```{r}
School2 <- School2 %>% drop_na(School.Name)
```

In the FARMS Percentage column, some of the values are represented by inequalities: <= 5.0 and >= 95.0. Since I only want numbers in this column, I'm going to replace these values with 5 and 95 respectively.

```{r}
School2[School2== ">= 95.0"] <-  "95.0"
```

```{r}
School2[School2== "<= 5.0"] <-  "5.0"
```

Now, I will convert the FARMS Percentage variable from characters to numerics.

```{r}
School2$FARMS.Pct <- as.numeric(School2$FARMS.Pct)
```

Taking a look at the NAs introduce by coercion.

```{r}
summary(School2)
```

For HLR_zip, I am renaming the columns.

```{r}
HLR_zip <- HLR_zip %>% 
  rename(Units = x) %>%
  rename(Property.Zip = Category) %>%
  rename(Ownership = Group.2)
```

#Perform descriptive statistics.

Let's find the mean graduation rates by county for all years combined (2015-2019).

```{r}
mean_grad <- drop_na(School2) %>%
select(LEA.Name, Grad.Rate) %>%
group_by(LEA.Name) %>%
summarise_all(funs(mean))
mean_grad
```
In preparation for visualizations, I am going to install the necessary packages.

```{r}
install.packages("ggplot2")
library("ggplot2")
install.packages("ggthemes")
library("ggthemes")
install.packages("highcharter")
library("highcharter")
```

I'm going to make a line graph of graduation rates by county over time.

```{r}
mean_grad2 <- drop_na(School2) %>%
select(LEA.Name, Grad.Rate, Academic.Year) %>%
group_by(Academic.Year, LEA.Name) %>%
summarise_if(is.numeric, mean)
mean_grad2
```
```{r}
mean_grad2$Grad.Rate <- as.numeric(mean_grad2$Grad.Rate)
```

```{r}
mean_grad2 %>% hchart(
  'line', hcaes(x = Academic.Year, y = Grad.Rate, group = LEA.Name))
```

From this graph, you can see that Montgomery County's graduation rates stay consistently higher than Prince George's County's, and both counties experiencing peaks and drops over the years, although not in the same pattern.

Turning to the HLR data, I want to find the top 10 zip codes in Montgomery County with the highest total units owned by the Housing Opportunities Community.

```{r}
HLR_zip %>% 
  subset(Ownership=='Housing Opportunities Community') %>%
  arrange(desc(Units))%>%
  head(10)
```
Visualize this on a bar graph.

```{r}
HOC_plot <- HLR_zip %>% 
  subset(Ownership=='Housing Opportunities Community') %>%
  arrange(desc(Units))%>%
  head(10) %>%
ggplot(HLR_zip, mapping = aes(x=Property.Zip, y=Units, color=Ownership)) +
  geom_bar(stat="identity", fill="orange") +
  theme(axis.text.x = element_text(angle = 45))
HOC_plot
```

These zip codes correspond to the following MCPS high schools: Bethesda-Chevy Chase, Walter Johnson, Northwest, Longview, Seneca Valley, Watkins Mill, Montgomery Blair, Northwood, John F. Kennedy, and Springbrook.

Now, I want to find the same data for Prince George's County. Since I don't have an equivalent dataset for PG County, I will use the low-income housing units data from HUD.

```{r}
PG_LIunits <- drop_na(School2) %>%
select(LEA.Name, Total_LIunits, School.Zip) %>%
subset(LEA.Name=="Prince George's") %>%
group_by(School.Zip) %>%
arrange(desc(Total_LIunits))%>%
head(70)
PG_LIunits
```

```{r}
PG_LI_plot <- 
ggplot(PG_LIunits, mapping = aes(x=School.Zip, y=Total_LIunits, color=LEA.Name)) +
  geom_bar(stat="identity", fill="purple") +
  theme(axis.text.x = element_text(angle = 45))
PG_LI_plot
```

These zip codes correspond to the following PGCPS high schools: Duval, Laurel, Chesapeake Math and IT Public Charter, Bladensburg, International High School @ Langley Park, Community Based Classrooms, C. Elizabeth Rieg Regional Center, Annapolis Road Academy, College Park Academy, Parkdale, Green Valley Academy at Edgar Allan Poe, Suitland, Central, Crossland, Crossland Evening/Saturday, Northwestern, Northwestern Evening/Saturday, and Fairmont Heights.

Now, I want to look at the FARMS data for both counties over time.

```{r}
mean_FARMS <- drop_na(School2) %>%
select(LEA.Name, FARMS.Pct, Academic.Year) %>%
group_by(Academic.Year, LEA.Name) %>%
summarise_if(is.numeric, mean)
mean_FARMS
```
```{r}
mean_FARMS$FARMS.Pct <- as.numeric(mean_FARMS$FARMS.Pct)
```

```{r}
mean_FARMS %>% hchart(
  'line', hcaes(x = Academic.Year, y = FARMS.Pct, group = LEA.Name))
```

This line graph shows us that Montgomery County's FARMS rates are consistently lower than Prince George's County's, and that the trends between counties appear to be the opposite of each other.

Next, I want to take a look at the MPDU data, and see which Planning Areas have the most and least MPDUs.

```{r}
MPDU_plot <- MPDU_Sums %>%
  hchart(
  'bar', hcaes(x = Planning.Area, y = Total.MPDU),
  color = "#49E3D0", borderColor = "black"
  )
MPDU_plot
```

Here we can see that Clarksburg has the most MPDUs, followed by "NA." It's a bit confusing that the "NA" planning area has so many MPDUs.

#Statistical analysis.

## FARMS and Graduation Rates.

I'm going to run a Pearson's Correlation Test on Graduation Rate and FARMS Percentage.

```{r}
Pearson_Grad_FARMS <- cor.test(School2$Grad.Rate, School2$FARMS.Pct, 
                    method = "pearson")
Pearson_Grad_FARMS
```

The results of this test suggest that these two variables have a statistically significant negative, moderate correlation with a correlation coefficient of -0.49.

I want to see if there's a significant difference between the average graduation rates between the two counties, so I'm giong to conduct a one-way ANOVA test.

```{r}
grad.aov <- aov(Grad.Rate ~ LEA.Name, data = School2)

summary(grad.aov)
```

Since the p-value is less than 0.05, we can conclude that there are significant differences between Graduation Rates across the two counties.

```{r}
hchart(School2, "scatter", hcaes(x = Grad.Rate, y = FARMS.Pct, group = LEA.Name))
```

```{r}
library("ggpubr")
ggscatter(School2, x = "Grad.Rate", y = "FARMS.Pct",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "Graduation Rates", ylab = "FARMS Percentage")
```

## FARMS and Low-Income Housing.

```{r}
Pearson_LI_FARMS <- cor.test(School2$FARMS.Pct, School2$Total_LIunits, 
                    method = "pearson")
Pearson_LI_FARMS
```

The p-value is less than 0.05, which suggests that amount of low-income units in the school's zip code and FARMS percentage of that school have a positive, weak correlation with a correlation coefficient of 0.23.

Here's a scatter plot of that data with a regression line.

```{r}
ggscatter(School2, x = "FARMS.Pct", y = "Total_LIunits",
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson",
          xlab = "FARMS Percentage", ylab = "Total Low-Income Housing Units")
```

I want to look at the relationships between all the continuous variables from my School data. To do this, I will make a correlation matrix/correlogram.

First, I will create a data frame that only includes the numeric variables from the School data frame.

```{r}
School_Numerics <- School2 %>% select(Grad.Rate, FARMS.Pct, Total_LIunits)
```

Now, I will make a correlation matrix out of my numerics-only data. I'll use the "cor" function for this and set "use" to "complete.obs" since my data has a lot of NAs.

```{r}
School_cor <- round(cor(School_Numerics, use = "complete.obs"), 2)
School_cor
```

Now for the correlogram.

```{r}
install.packages("corrplot")
library("corrplot")
```

```{r}
corrplot(School_cor, method = "circle")
```

This plot makes it easier for us to see that out of these three variables, FARMS Percentage and Graduation Rate have the strongest correlation, followed by FARMS Percentage and Total Low-Income Housing Units, and lastly Graduation Rates and Total Low-Income Housing Units, which had the weakest correlation.
