---
title: "DATA205 EDA"
output: html_document
---

## Load data
### Read in two dataMontgomery datasets: "Moderately Priced Dwelling Units" and "Housing Licensing and Registration"
### Install necessary packages for data cleaning, wrangling, and visualizing: tidyverse and ggplot2.

```{r}
MPDU <- read.csv("https://data.montgomerycountymd.gov/api/views/pbh5-g4e5/rows.csv?accessType=DOWNLOAD")
HLR <- read.csv("https://data.montgomerycountymd.gov/api/views/et5s-xste/rows.csv?accessType=DOWNLOAD")
install.packages("tidyverse")
install.packages("ggplot2")
library("tidyverse")
library("ggplot2")
```


## Look at the structure of the MPDU dataset.

```{r}
str(MPDU)
```

### Look at the head and tail of the MPDU dataset.

```{r}
head(MPDU)
tail(MPDU)
```

### Create new versions of the datasets without NAs.

```{r}
MPDU2 <- drop_na(MPDU)
HLR2 <- drop_na(HLR)
```


### Take a look at the summary of the MPDU dataset.

```{r}
summary(MPDU)
```


### Look at the structure, head, and tail of the HLR dataset.

```{r}
str(HLR)
head(HLR)
tail(HLR)
```




### Take a look at the summary of the HLR dataset.

```{r}
summary(HLR)
```


```{r}
MPDU$Total.Units <- as.numeric(MPDU$Total.Units)
MPDU$Total.MPDU <- as.numeric(MPDU$Total.MPDU)
MPDU3 <- MPDU %>% 
select(Planning.Area, Total.Units, Total.MPDU) %>%
group_by(Planning.Area)
MPDU3
```
```{r}
MPDU4 <- drop_na(MPDU3)
```

```{r}
MPDU5 <- MPDU4 %>% 
    summarize_if(is.numeric, sum)
MPDU5
```

