---
title: "DATA-205-Capstone-Project"
author: "Fiona Caretto"
date: "4/28/2021"
output: html_document
---

# Loading In Datasets

First, I read in two dataMontgomery datasets: "Moderately Priced Dwelling Units" (named as "MPDU") and "Housing Licensing and Registration" (named as "HLR").

```{r}
MPDU <- read.csv("https://data.montgomerycountymd.gov/api/views/pbh5-g4e5/rows.csv?accessType=DOWNLOAD")
HLR <- read.csv("https://data.montgomerycountymd.gov/api/views/et5s-xste/rows.csv?accessType=DOWNLOAD")
```

Next, I will read in the "Students Receiving Special Services" dataset from the Maryland State Department of Education, one for each year 2015-2019. Each of these will be called "SS+last two digits of year."

```{r}
url1 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/378"
download.file(url1, "Special_Services_2019.zip")
unzip("Special_Services_2019.zip")
SS19 <- read.csv("Special_Services_2019.csv")
```

```{r}
url2 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/341"
download.file(url2, "Special_Services_2018.zip")
unzip("Special_Services_2018.zip")
SS18 <- read.csv("Special_Services_2018.csv")
```

```{r}
url3 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/322"
download.file(url3, "Special_Services_2017.zip")
unzip("Special_Services_2017.zip")
SS17 <- read.csv("Special_Services_2017.csv")
```

```{r}
url4 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/298"
download.file(url4, "Special_Services_2016.zip")
unzip("Special_Services_2016.zip")
SS16 <- read.csv("Special_Services_2016.csv")
```

```{r}
url5 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/258"
download.file(url5, "Special_Services_2015.zip")
unzip("Special_Services_2015.zip")
SS15 <- read.csv("Special_Services_2015.csv")
```

Now, I will read in the "Adjusted Cohort Graduation Rate" datasets also from the Maryland State Department of Education, also for 2015-2019. Each of these will be named "Grad+last two digits of year."

```{r}
url6 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/390"
download.file(url6, "Cohort_Graduation_and_Dropout_Rates_2019.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2019.zip")
Grad19 <- read.csv("Cohort_Grad_Rate_2019.csv")
```

```{r}
url7 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/354"
download.file(url7, "Cohort_Graduation_and_Dropout_Rates_2018.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2018.zip")
Grad18 <- read.csv("Cohort_Grad_Rate_2018.csv")
```

```{r}
url8 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/326"
download.file(url8, "Cohort_Grad_Rate_2017.zip")
unzip("Cohort_Grad_Rate_2017.zip")
Grad17 <- read.csv("Cohort_Grad_Rate_2017.csv")
```

```{r}
url9 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/302"
download.file(url9, "Cohort_Grad_Rate_2016.zip")
unzip("Cohort_Grad_Rate_2016.zip")
Grad16 <- read.csv("Cohort_Grad_Rate_2016.csv")
```

```{r}
url10 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/262"
download.file(url10, "Cohort_Grad_Rate_2015.zip")
unzip("Cohort_Grad_Rate_2015.zip")
Grad15 <- read.csv("Cohort_Grad_Rate_2015.csv")
```
