---
title: "DATA205 EDA"
output: html_document
---

# Exploratory Data Analysis (EDA)

## Begin to load in the data

### Read in two dataMontgomery datasets: "Moderately Priced Dwelling Units" (named as "MPDU") and "Housing Licensing and Registration" (named as "HLR")

### Install necessary packages for data cleaning, wrangling, and visualizing: tidyverse and ggplot2.

```{r}
MPDU <- read.csv("https://data.montgomerycountymd.gov/api/views/pbh5-g4e5/rows.csv?accessType=DOWNLOAD")
HLR <- read.csv("https://data.montgomerycountymd.gov/api/views/et5s-xste/rows.csv?accessType=DOWNLOAD")
install.packages("tidyverse")
install.packages("ggplot2")
library("tidyverse")
library("ggplot2")
```


## Look at the structure of the MPDU dataset.

```{r}
str(MPDU)
```

### Look at the head and tail of the MPDU dataset.

```{r}
head(MPDU)
tail(MPDU)
```

### Create new versions of the datasets without NAs.

```{r}
MPDU2 <- drop_na(MPDU)
HLR2 <- drop_na(HLR)
```

### Actually, I don't want to do this now so I will delete these two new dataframes.

```{r}
remove(MPDU2)
remove(HLR2)
```


### Take a look at the summary of the MPDU dataset.

```{r}
summary(MPDU)
```


### Look at the structure, head, and tail of the HLR dataset.

```{r}
str(HLR)
head(HLR)
tail(HLR)
```



### Take a look at the summary of the HLR dataset.

```{r}
summary(HLR)
```


## Create a new data frame for the MPDU dataset showing just Total Units and Total MPDU, grouped by Planning Area.

```{r}
MPDU$Total.Units <- as.numeric(MPDU$Total.Units)
MPDU$Total.MPDU <- as.numeric(MPDU$Total.MPDU)
MPDU_Units_MPDU_PA <- MPDU %>% 
select(Planning.Area, Total.Units, Total.MPDU) %>%
group_by(Planning.Area)
MPDU_Units_MPDU_PA
```
### Remove unused dataframe.

```{r}
remove(MPDU3)
```


### Drop the NAs from this new data frame.

```{r}
New_MPDU_Units_MPDU_PA <- drop_na(MPDU_Units_MPDU_PA)
```

### Create a new data frame showing the sums of Total Unit and Total MPDU columns grouped by Planning Area.

```{r}
MPDU_Sums_PA <- New_MPDU_Units_MPDU_PA %>% 
    summarize_if(is.numeric, sum)
MPDU_Sums_PA
```

### Remove more unneeded data frames.

```{r}
remove(MPDU4)
remove(MPDU5)
```


### Create a new data frame with variables of interest from the HLR dataset: Ownership Type, Property City, Structure Type, Unit Count, and License Type.

```{r}
HLR_New <- HLR %>% 
select(Ownership.Type, Property.City, Property.Zip.Code, Structure.Type, Unit.Count, License.Type) %>%
group_by(Property.City)
HLR_New
```
### Remove unused data frame.

```{r}
remove(HLR3)
```


## Read in Students Receiving Special Services 2019 dataset.

```{r}
url2 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/378"
download.file(url2, "Special_Services_2019.zip")
unzip("Special_Services_2019.zip")
SS19 <- read.csv("Special_Services_2019.csv")
```

## Read in Students Receiving Special Services 2018 dataset.

```{r}
url5 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/341"
download.file(url5, "Special_Services_2018.zip")
unzip("Special_Services_2018.zip")
SS18 <- read.csv("Special_Services_2018.csv")
```


## Read in Students Receiving Special Services 2017 dataset.

```{r}
url6 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/322"
download.file(url6, "Special_Services_2017.zip")
unzip("Special_Services_2017.zip")
SS17 <- read.csv("Special_Services_2017.csv")
```


## Read in Students Receiving Special Services 2016 dataset.

```{r}
url7 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/298"
download.file(url7, "Special_Services_2016.zip")
unzip("Special_Services_2016.zip")
SS16 <- read.csv("Special_Services_2016.csv")
```


## Read in Students Receiving Special Services 2015 dataset.

```{r}
url8 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/258"
download.file(url8, "Special_Services_2015.zip")
unzip("Special_Services_2015.zip")
SS15 <- read.csv("Special_Services_2015.csv")
```


## Read in Cohort Graduation Rates 2019 dataset.

```{r}
url3 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/390"
download.file(url3, "Cohort_Graduation_and_Dropout_Rates_2019.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2019.zip")
Grad19 <- read.csv("Cohort_Grad_Rate_2019.csv")
```

## Read in Cohort Graduation Rates 2018 dataset.

```{r}
url9 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/354"
download.file(url9, "Cohort_Graduation_and_Dropout_Rates_2018.zip")
unzip("Cohort_Graduation_and_Dropout_Rates_2018.zip")
Grad18 <- read.csv("Cohort_Grad_Rate_2018.csv")
```


## Read in Cohort Graduation Rates 2017 dataset.

```{r}
url10 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/326"
download.file(url10, "Cohort_Grad_Rate_2017.zip")
unzip("Cohort_Grad_Rate_2017.zip")
Grad17 <- read.csv("Cohort_Grad_Rate_2017.csv")
```


## Read in Cohort Graduation Rates 2016 dataset.

```{r}
url11 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/302"
download.file(url11, "Cohort_Grad_Rate_2016.zip")
unzip("Cohort_Grad_Rate_2016.zip")
Grad16 <- read.csv("Cohort_Grad_Rate_2016.csv")
```


## Read in Cohort Graduation Rates 2015 dataset.

```{r}
url12 <- "https://reportcard.msde.maryland.gov/DataDownloads/FileDownload/262"
download.file(url12, "Cohort_Grad_Rate_2015.zip")
unzip("Cohort_Grad_Rate_2015.zip")
Grad15 <- read.csv("Cohort_Grad_Rate_2015.csv")
```


## Read in Low-Income Housing Tax Credit dataset.

```{r}
url4 <- "https://lihtc.huduser.gov/lihtcpub.zip"
download.file(url4, "lihtcpub.zip")
unzip("lihtcpub.zip")
LIHTC <- read.csv("LIHTCPUB.csv")
```


## Create two data frames: one for MCPS and one for PGCPS with the names of every High School and their corresponding zip code.

```{r}
MCPS <- data.frame(
  MCPS_name = c("Stephen Knolls School", "Longview School", "Montgomery Blair High", "Northwest High", "Walter Johnson High", "Richard Montgomery High", "Gaithersburg High", "Clarksburg High", "Winston Churchill High", "Bethesda-Chevy Chase High", "Quince Orchard High", "Thomas S. Wootton High", "Walt Whitman High", "Wheaton High", "Sherwood High", "Paint Branch High", "John F. Kennedy High", "Albert Einstein High", "Northwood High", "Springbrook High", "James Hubert Blake High", "Watkins Mill High", "Col. Zadok Magruder High", "Rockville High", "Damascus High", "Poolesville High", "Seneca Valley High", "Alternative Programs", "John L Gildner Regional Inst for Children & Adol", "Rock Terrace School"),
  MCPS_zip = c(20895, 20874, 20901, 20874, 20814, 20852, 20877, 20871, 20854, 20814, 20878, 20850, 20817, 20906, 20860, 20866, 20902, 20895, 20901, 20904, 20905, 20879, 20855, 20851, 20872, 20837, 20874, 20853, 20850, 20850)
)
PGCPS = data.frame(
  PGCPS_name = c("Eleanor Roosevelt High","High Point High","Bowie High","Parkdale High","Northwestern High","Dr. Henry A. Wise, Jr. High","Charles Herbert Flowers High","Duval High","Suitland High","Bladensburg High","Laurel High","Oxon Hill High","Potomac High","Frederick Douglass High","Crossland High","Gwynn Park High","Largo High","Central High","Friendly High","Fairmont Heights High","Surrattsville High","Academy of Health Sciences at PGCC","Chesapeake Math and IT Public Charter","International High school @ Largo","Northwestern Evening/Saturday High","International High School @ Langley Park","Chesapeake Math and IT South Public Charter","College Park Academy","Crossland Evening/Saturday High","Tall Oaks High","Croom High","Community Based Classrooms","C. Elizabeth Rieg Regional Center","James E. Duckworth Regional Center"),
  PGCPS_zip = c(20770, 20705, 20715, 20737, 20782, 20772, 20774, 20706, 20746, 20710, 20707, 20745, 20745, 20772, 20748, 20613, 20774, 20747, 20744, 20785, 20735, 20774, 20707, 20772, 20782, 20710, 20772, 20737, 20748, 20721, 20623, 20710, 20716, 20705)
)
MCPS
PGCPS
```

## Rename the Class.of.Year column in Grad15 dataset to match the rest of the Graduation Rate datasets.

```{r}
Grad15 <- Grad15 %>% 
  rename(Academic.Year = Class.of.Year)
```

## Rename the Class.of.Year column in Grad16 dataset to match the rest of the Graduation Rate datasets.

```{r}
Grad16 <- Grad16 %>% 
  rename(Academic.Year = Class.of.Year)
```


## Rename more columns in Graduation Rates datasets to match the rest of them.

```{r}
Grad17 <- Grad17 %>%
  rename(LEA = LEA.Number)
Grad18 <- Grad18 %>%
  rename(LEA = LSS.Number, LEA.Name = LSS.Name)
Grad19 <- Grad19 %>%
  rename(LEA = LSS.Number, LEA.Name = LSS.Name)
```

## Combine all Graduation Rates datasets into one Total Graduation Rate dataset.

```{r}
Total_Grad <- rbind(Grad15, Grad16, Grad17, Grad18, Grad19)
```

## Rename some columns to match the other Students Receiving Special Services datasets and then create new datasets pulling in only the variables I am interested in.

```{r}
SS15_2 <- SS15 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct, FARMS.Cnt, Enrollment.Cnt)
SS15_2

SS16_2 <- SS16 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct, FARMS.Cnt, Enrollment.Cnt)
SS16_2

SS17_2 <- SS17 %>% 
select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct, FARMS.Cnt, Enrollment.Cnt)
SS17_2

SS18_2 <- SS18 %>% 
  rename(LEA.Name = LSS.Name) %>%
  select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct, FARMS.Cnt, Enrollment.Cnt)
SS18_2

SS19_2 <- SS19 %>% 
  rename(LEA.Name = LSS.Name) %>%
  select(Academic.Year, LEA.Name, School.Name, School.Type, FARMS.Pct, FARMS.Cnt, Enrollment.Cnt)
SS19_2
```

## Combine all Students Receiving Special Services datasets into one Total Special Services dataset.

```{r}
Total_SS <- rbind(SS15_2, SS16_2, SS17_2, SS18_2, SS19_2)
```


## Count NAs for Total_SS and Total_Grad datasets.

```{r}
summary(Total_Grad)
```

```{r}
summary(Total_SS)
```

### Great, there are no NAs that need to be removed.

## Create a new Total_Grad dataset with only Montgomery and Prince George's counties.

```{r}
Total_Grad_MCPS_PGCPS <- subset(Total_Grad, LEA.Name == "Montgomery" & Cohort == "4-year adjusted cohort" | LEA.Name == "Prince George's" & Cohort == "4-year adjusted cohort")
```


### Convert the Grad.Rate variable into a numeric value.

```{r}
Total_Grad_MCPS_PGCPS$Grad.Rate <- as.numeric(Total_Grad_MCPS_PGCPS$Grad.Rate)
```

### Take a look at the NAs that were introduced after converting the Grad.Rate variable into a numeric value.


```{r}
summary(Total_Grad_MCPS_PGCPS)
```


### Drop the NAs from the Total_Grad_MCPS_PGCPS data frame and then make a new data frame showing the mean graduation rates by county for all years combined (2015-2019).

```{r}
mean_grad <- drop_na(Total_Grad_MCPS_PGCPS) %>%
select(LEA.Name, Grad.Rate) %>%
group_by(LEA.Name) %>%
summarise_all(funs(mean))
mean_grad
```
### Make a boxplot for graduation rates by county from 2015-2019. Not a great visualization.

```{r}
plot1<- ggplot(data=Total_Grad_MCPS_PGCPS, aes(x=Academic.Year, y=Grad.Rate, group=LEA.Name)) +
  geom_boxplot()
plot1
```

### Look at the summary of our new HLR dataset.

```{r}
summary(HLR_New)
```


### Convert the Property.Zip.Code variable into a numeric value.

```{r}
HLR_New$Property.Zip.Code <- as.numeric(HLR_New$Property.Zip.Code)
```

### Look again at the summary of the HLR dataset to observe new NAs that were added due to the conversion of the Property.Zip.Code variable into a numeric value.

```{r}
summary(HLR_New)
```


### Create a new data frame from the HLR dataset showing the total amount of units under each Ownership Type by Property Zip Code.

```{r}
HOC_zip <- aggregate(HLR_New$Unit.Count, by=list(Category=HLR_New$Property.Zip.Code, HLR_New$Ownership.Type), FUN=sum)
HOC_zip
```

### Find the top 10 zip codes in Montgomery County with the highest total units owned by the Housing Opportunities Community.

```{r}
HOC_zip %>% 
  subset(Group.2=='Housing Opportunities Community') %>%
  arrange(desc(x))%>%
  head(10)
```
### Visualize this data on a bar plot, showing total amount of units owner by the HOC on the y-axis and zip code on the x-axis. Not a great graphic.

```{r}
plot2 <- HOC_zip %>% 
  subset(Group.2=='Housing Opportunities Community') %>%
  arrange(desc(x))%>%
  head(10)%>%
ggplot(HOC_zip, mapping = aes(x=Category, y=x, color=Group.2)) +
  geom_bar(stat="identity", fill="white")
plot2
```

### Take a look at the summary of LIHTC.

```{r}
summary(LIHTC)
```

### Take a look at the summary of Total_SS.

```{r}
summary(Total_SS)
```

### Convert the fips2010 variable from the LIHTC dataset into a numeric value. This variable is the FIPS code of the public housing project, which contains the State FIPS code, County FIPS code, and Census Tract number.

```{r}
LIHTC$fips2010 <- as.numeric(LIHTC$fips2010)
LIHTC_county <- LIHTC %>%
  filter((substr(fips2010, 1, 5) == "24031") | (substr(fips2010, 1, 5) == "24033"))
```

### Find the zip codes in Prince George's County with the top 10 most Low Income Housing Units.

```{r}
LIHTC_PG_zip <- LIHTC_county %>%
  filter(substr(fips2010, 1, 5) == "24033") %>%
  select(latitude, longitude, proj_zip, fips2010, li_units) %>%
  arrange(desc(li_units))%>%
  head(20)
LIHTC_PG_zip
```
## The top 10 zip codes in PG county with the most Low Income Housing Units are: 20747, 20785, 20746, 20710, 20706, 20783, 20712, 20737, 20782, 20708. These correlate to the following PGCPS High Schools: Central High School, Fairmont Heights High School, Suitland High School, Bladensburg High School, International High School @ Langley Park, Community Based Classrooms, Duval High School, Parkdale High School, College Park Academy, Northwestern High School, Northwestern Evening/Saturday High School.


### Check the class type of the Graduation Rates data frame.

```{r}
class(Total_Grad_MCPS_PGCPS)
```
### Install and load data.table

```{r}
install.packages("data.table")           
library("data.table")
```


### Create a table out of the Graduation Rates data frame.

```{r}
Total_Grad_Table <- setDT(Total_Grad_MCPS_PGCPS)
class(Total_Grad_Table)
```

### Merge the Graduation Rates data frame and Students Receiving Special Services data frame into a full data frame of all the school data.

```{r}
All_Grad_SS <- merge(Total_Grad_MCPS_PGCPS, Total_SS)
```


### Remove the rows for "All (MCPS/PGCPS) Schools" from the data frame.
#### These values alter my overall data.

```{r}
All_Grad_SS <- All_Grad_SS[-c(2, 3, 4, 44, 45, 46), ]
```

### Filter out Elementary and Middle Schools so we only have High Schools left in our data frame.

```{r}
All_Grad_SS <- All_Grad_SS %>%
  filter(substr(School.Type, 1, 4) == "High")
```


